---
title: "Assignment 1"
author: Xuan Gong
date: March 13, 2017
output: md_document
---

#Assignment 1 for Reproducible Research Class

===================================================

## Loading and preprocessing the data

1. load the data
```{r loaddata, echo=TRUE, results="hide"}
df <- read.csv(file = "activity.csv", na.strings = "NA")
```

## What is mean total number of steps taken per day?

1. Calculate the total number of steps taken per day
```{r totalstep, echo=TRUE}
total.steps <- with(df, tapply(steps, date, sum, na.rm=T))
```

2. Make a histogram of the total number of steps taken each day
```{r plot01, results="asis"}
hist(total.steps, main = "Histogram of Total Steps per Day", xlab = "Total Steps per Day")
```

3. Calculate and report the mean and median of the total number of steps taken per day

Mean:
```{r}
mean(total.steps)
```

Median:
```{r}
median(total.steps)
```

##What is the average daily activity pattern?

1. calculate the average number of steps by 5-minute interval
```{r avestep, echo=TRUE}
average.steps <- with(df, tapply(steps, interval, mean, na.rm=T))
```

2. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r plot02, echo=TRUE, results="asis"}
plot(names(average.steps), average.steps, type = "l", main = "Plot of Average Steps per Interval", xlab = "Interval", ylab = "Average Steps across all days")
```

3. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
names(which.max(average.steps))
```


##Imputing missing values

1. Calculate and report the total number of missing values in the dataset
```{r}
sum(is.na(df))
```

2. filling in all of the missing values in the dataset with the mean for that interval, and create a new dataset
```{r}
na <- is.na(df$steps)
df.na <- df[na,]
df.non.na <- df[!na,]
x <- 0
for (x in unique(df.na$interval)) {
        df.na[which(df.na$interval == x), 1] <- average.steps[which(names(average.steps) == x)]
        x <- x + 5
}
dfnew <- rbind(df.na, df.non.na)
```

3. Make a histogram of the total number of steps taken each day
```{r plot3, echo=TRUE}
total.steps.new <- with(dfnew, tapply(steps, date, sum, na.rm=T))
hist(total.steps.new, main = "Histogram of Total Steps per Day", xlab = "Total Steps per Day")
```

4. Calculate and report the mean and median total number of steps taken per day.

Mean:
```{r}
mean(total.steps.new)
```

Median:
```{r}
median(total.steps.new)
```

Do these values differ from the estimates from the first part of the assignment? 

Yes.

What is the impact of imputing missing data on the estimates of the total daily number of steps?

Imputing missing data increased the frequency of range 10000 ~ 15000 steps per day.

##Are there differences in activity patterns between weekdays and weekends?

1. Create a new factor variable in the dataset with two levels ¨C ¡°weekday¡± and ¡°weekend¡± indicating whether a given date is a weekday or weekend day.

```{r}
library(lubridate)
library(dplyr)
dfnew$date <- as.Date(as.character(df$date), "%Y-%m-%d")
dfnew.addfactor <- mutate(dfnew, weekday = factor(wday(date) %in% c(1,7), labels = c("weekday", "weekend")))
```

2. Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).

```{r plot4, echo=TRUE}
average.steps.new <- with(dfnew.addfactor, tapply(steps, list(interval, weekday), mean, na.rm=T))
type <- unique(dfnew.addfactor$weekday)
library(reshape2)
dfplot <- melt(average.steps.new, measure.vars = type)
library(ggplot2)
g <- ggplot(dfplot, aes(Var1, value))
g + geom_line(aes(colour = Var2)) + facet_grid(Var2 ~ .) + labs(title = "Plot of Average Steps per Interval") + labs(x = "Interval") + labs(y = "Number of Steps")
```

